Model of lens candidates from Space Warps CFHTLS

### 2016-07-11 paper2.tex

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[R. Kueng et al.]{
Rafael Kueng,$^{1}$\thanks{E-mail: rafael.kueng@uzh.ch}
Prasenjit Saha,$^{1}$
Lucy Oswald$^{3}$
%and Fourth Author$^{4}$
\\
% List of institutions
$^{1}$Physik Institut, University of Zurich, Winterthurerstrasse XXX, Zurich 8057, Switzerland\\
$^{2}$Department, Institution, Street Address, City Postal Code, Country\\
$^{3}$Another Department, Different Institution, Street Address, City Postal Code, Country
%$^{4}$Another Department, Different Institution, Street Address, City Postal Code, Country
}

@@

While SpaghettiLens is citizen science and based on Space~Warps, the
collaboration with volunteers takes a very different form.  In
Space~Warps, volunteers in a crowd of $\gtrsim10^4$ make independent
contributions.  Each person is presented with a random selection of
survey-patches and invited to (in effect) vote on each.  The system
estimates each volunteer's skill level according to test-patches
interspersed with the real data, and weights their votes accordingly
\citep{2016MNRAS.455.1171M}.  There is an active forum for volunteers,
but since everyone is seeing different data samples with minimal
overlap, the forum has little if any influence on votes.  With
SpaghettiLens there is a small group of volunteers and their
contributions are interdependent.\footnote{For more on how this works,
  see ``Collaborative gravitational lens modelling\dots'' in {\tt
    http://letters.zooniverse.org} especially the model tree.}

Section~\ref{sec:spl} explains the SpaghettiLens system.

@@

@@

\subsection{Parameterisation of pixel models} \label{sec:parameter}
In order to fit the set of pixelated models to a single parameterised model, a program was written that took a parameterised function and subtracted from it the mean and the principle components of the data, which were calculated using classical Principle Component Analysis.
This created the residuals function.
The number of components defined as principle was varied to test how this affected the output, and it was found that using 5 principle components tended to give a reasonable approximation.
A masking function was added which selected only the data points that fell inside the image of the lens, and the principal components were clipped in order to keep the values inside the region of the ensemble of models.
Any value higher than the clip was set to be the clip value.
This was chosen to be 2.5 as, assuming that the data follows a Gaussian error distribution, almost all the values for the variance should lie between 2 and 3 standard deviations from the mean.
Minimising the residuals function produces the set of parameters that fit the parameterised function to the original pixelated ensemble most closely.
A least squares fit was used to perform this minimisation.
The parameterised model function was obtained from the gravitational potential of an isothermal ellipsoid mass distribution \cite{2001astro.ph..2341K}.
This model is frequently used to describe gravitational lenses as it tends to fit well with observations.
The isothermal ellipsoid model outputs three useful parameters: the radius of the Einstein ring, the ellipticity of the model and the angle of the ellipticity from the vertical, giving the orientation of the galaxy.
By applying this model to simulated lenses for which the values of these parameters were already known, it was possible to gain an estimate of the projected accuracy of the results, before applying the model to the candidate lensing galaxies.

@@

Based on \cite{2003MNRAS.344.1000B} models.

>>>>

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[R. K\"ung et al.]{
Rafael K\"ung,$^{1}$\thanks{E-mail: rafael.kueng@uzh.ch}
Prasenjit Saha,$^{1}$
Lucy Oswald$^{3}$
%and Fourth Author$^{4}$
\\
% List of institutions
$^{1}$Physik Institut, University of Zurich, Winterthurerstrasse XXX, Zurich 8057, Switzerland\\
$^{2}$Department, Institution, Street Address, City Postal Code, Country\\
$^{3}$Another Department, Different Institution, Street Address, City Postal Code, Country
%$^{4}$Another Department, Different Institution, Street Address, City Postal Code, Country
}

@@

While SpaghettiLens is citizen science and based on Space~Warps, the
collaboration with volunteers takes a very different form.  In
Space~Warps, volunteers in a crowd of $\gtrsim10^4$ make independent
contributions.  Each person is presented with a random selection of
survey-patches and invited to (in effect) vote on each.  The system
estimates each volunteer's skill level according to test-patches
interspersed with the real data, and weights their votes accordingly
\citep{2016MNRAS.455.1171M}.  There is an active forum for volunteers,
but since everyone is seeing different data samples with minimal
overlap, the forum has little if any influence on votes.  With
SpaghettiLens there is a small group of volunteers and their
contributions are interdependent.\footnote{See ``Collaborative
  gravitational lens modelling\dots'' in {\tt
    http://letters.zooniverse.org} especially the model tree.}

Section~\ref{sec:spl} summarises the basics of the SpaghettiLens
system and discusses its ongoing development.  Basically, the system
delivers free-form reconstructions (with uncertainties) of the
sky-projected mass distributions of lens candidates.  These represent
a consensus among modelling volunteers as to the best that could be
achieved with the available data and software.  Photometric lens
redshifts are used; source redshifts are set to $z=2$.  The models
can, however, be trivially rescaled to use better redshift values, as
and when they become available.

@@

\cite{2015MNRAS.447.2170K} introduced and implemented two
main ideas. One is a `spaghetti diagram,' which is a cartoon-like
markup of a candidate lensing system which encodes a proposal for its
mass distribution.  It enables volunteers to interact with a
server-side numerical engine \citep[GLASS, developed
  by][]{2014MNRAS.445.2181C} in an intuitive way.  The other new idea
was a `model tree' --- a volunteer can model a lens from scratch, or
they can modify an existing model and thus try to improve upon it.
This facility, in conjunction with forum discussion, enables
development trees of models.  The models reported in this paper are
the most recent ones for each lens candidate, but typically they are
descended from previous models through the contributions of more than
one individual.

@@

\subsection{Parameterisation of pixel models} \label{sec:parameter}
\todo{}
In order to fit the set of pixelated models to a single parameterised model, a program was written that took a parameterised function and subtracted from it the mean and the principle components of the data, which were calculated using classical Principle Component Analysis.
This created the residuals function.
The number of components defined as principle was varied to test how this affected the output, and it was found that using 5 principle components tended to give a reasonable approximation.
A masking function was added which selected only the data points that fell inside the image of the lens, and the principal components were clipped in order to keep the values inside the region of the ensemble of models.
Any value higher than the clip was set to be the clip value.
This was chosen to be 2.5 as, assuming that the data follows a Gaussian error distribution, almost all the values for the variance should lie between 2 and 3 standard deviations from the mean.
Minimising the residuals function produces the set of parameters that fit the parameterised function to the original pixelated ensemble most closely.
A least squares fit was used to perform this minimisation.
The parameterised model function was obtained from the gravitational potential of an isothermal ellipsoid mass distribution \cite{2001astro.ph..2341K}.
This model is frequently used to describe gravitational lenses as it tends to fit well with observations.
The isothermal ellipsoid model outputs three useful parameters: the radius of the Einstein ring, the ellipticity of the model and the angle of the ellipticity from the vertical, giving the orientation of the galaxy.
By applying this model to simulated lenses for which the values of these parameters were already known, it was possible to gain an estimate of the projected accuracy of the results, before applying the model to the candidate lensing galaxies.

@@

Based on \cite{2003MNRAS.344.1000B} models.\todo{Needs text.}
